---
title: "Cherry Blossoms Prediction"
author: "Nafeez Chowdhury"
date: "2/28/2023"
output: html_document
---
Model

```{r,echo=T}
library(knitr)
library(tidyverse)
library(rnoaa)
library(rnpn)
library(rvest)
library(Hmisc)
library(caret)
library(lubridate)
theme_set(theme_bw(base_size = 20))
```

Washington, DC 2023
```{r,echo=T}
washingtondc <- read.csv('/Users/nafeezchowdhury/Documents/STAT 490/washingtondc.csv',header=T)
washingtondc1 <- washingtondc %>% filter(year >= 1980)

dc_temp <- ghcnd_search(stationid = "USC00186350",
                     var = c("tmax", "tmin"),
                     date_min = "1980-01-01",
                     date_max = "2022-12-31") %>%
  reduce(left_join) %>%
  transmute(year = parse_number(format(date, "%Y")),
            date, tmax = tmax / 10, tmin = tmin / 10,
            temp = (tmax + tmin) / 2)
dc_temp %>% slice(1:4) %>% select(year, date, temp) %>%
  kable(align = "c", col.names =
          c("year", "date", "midrange temperature °C"))

#########
dc_year_temps <-
  dc_temp %>% 
  mutate(doy_start = parse_number(format(date, "%j"))) %>%
  group_by(year) %>% 
  nest() %>%
  mutate(jan_temp = map(data, ~ mean(.$temp[1:30], na.rm = TRUE)),
         feb_temp = map(data, ~ mean(.$temp[31:60], na.rm = TRUE)),
         mar_temp = map(data, ~ mean(.$temp[61:90], na.rm = TRUE))) %>% 
  unnest(jan_temp) %>% 
  unnest(feb_temp) %>%
  unnest(mar_temp) %>%
  ungroup() %>%
  mutate(`very cold` = jan_temp < 4.6 & feb_temp < 4.6,
         `cold` = jan_temp < 4.6, feb_diff = feb_temp - jan_temp, mar_diff = mar_temp - jan_temp,
         `temp level` = factor(c(ifelse(`cold` == T, "cold", "warm"))), 
         grp = factor(c(ifelse(mar_diff < 4.7, "g1",ifelse(mar_diff < 7.2,'g2','g3')))))
#########

dc_sum_temps <- dc_temp %>% na.omit() %>% select(year,date,temp) %>% group_by(year) %>% 
  filter(1 <= yday(date) & yday(date) <= 100) %>%
  mutate(threshold = cumsum(temp))
dc_sum_temps1 <- dc_sum_temps %>% group_by(year) %>% nest() %>% left_join(washingtondc1) %>%
  ungroup()

vline_data <- data.frame(year = dc_sum_temps1$year, vline = dc_sum_temps1$bloom_doy)
curve_data <- dc_sum_temps1 %>% pull(data) %>% reduce(bind_rows) %>% mutate(year = format(date, "%Y"))

dc_csum_plot <- ggplot() + 
  geom_smooth(aes(x=yday(date),y=threshold,color=), data = curve_data) +
  geom_vline(aes(xintercept = vline), data = vline_data, linetype = 2) +
  facet_wrap(~year) + 
  theme_bw() +
  geom_smooth(method = "lm", aes(x=yday(date),y=threshold), 
              data = curve_data %>% mutate(doy = parse_number(format(date, "%j"))) %>% filter(doy >= 75), 
              color = "red") + coord_cartesian(ylim=c(0,800))
dc_csum_plot
#########

curve_data$doy <- parse_number(format(curve_data$date, "%j")) - 75
reg_line <- lm(threshold ~ yday(date):year, curve_data, subset = doy >= 0)$coefficients
threshold_100_days <- reg_line[1] + reg_line[-1]*100

washingtondc2 <- washingtondc1[-23,] %>% mutate(threshold = threshold_100_days)
washingtondc2$`temp level` <- dc_year_temps$`temp level`[-23]
mod <- lm(washingtondc2$bloom_doy~threshold_100_days)
summary(mod)
dcplot <- ggplot(washingtondc2,aes(x=threshold,y=bloom_doy,color=`temp level`)) +
  geom_point() + theme_bw() +
  geom_text(aes(label=year),hjust=0, vjust=0)
dcplot
########

dc_year_temps[-23,] %>% ggplot(aes(x=feb_diff,y=mar_diff,color=grp)) + geom_point() + 
  geom_text(aes(label=year,hjust=-0.5, vjust=0)) +
  geom_text(aes(label=washingtondc2$bloom_doy),hjust=1.3, vjust=0) +
  geom_text(aes(label=signif(washingtondc2$threshold,3)),hjust=1.3, vjust=1.3)

dc_g1 <- dc_year_temps[-23,] %>% mutate(bloom_doy = washingtondc2$bloom_doy, threshold = washingtondc2$threshold) %>% 
  filter(grp == 'g1')
dc_g2 <- dc_year_temps[-23,] %>% mutate(bloom_doy = washingtondc2$bloom_doy, threshold = washingtondc2$threshold) %>% 
  filter(grp == 'g2')
dc_g3 <- dc_year_temps[-23,] %>% mutate(bloom_doy = washingtondc2$bloom_doy, threshold = washingtondc2$threshold) %>% 
  filter(grp == 'g3')

washingtondc2 %>% ggplot(aes(x=threshold,y=bloom_doy,color=dc_year_temps$grp[-23])) + 
  geom_smooth(method='lm') + labs(y='bloom day of year',color='Group')

dc_temp_g1 <- lm(bloom_doy~threshold,data=dc_g1[-c(1,4),])
summary(dc_temp_g1)
plot(bloom_doy~threshold,data=dc_g1[-c(1,4),])
dc_temp_g2 <- lm(bloom_doy~threshold,data=dc_g2[-3,])
summary(dc_temp_g2)
plot(bloom_doy~threshold,data=dc_g2[-3,])
dc_temp_g3 <- lm(bloom_doy~threshold,data=dc_g3[-2,])
summary(dc_temp_g3)
plot(bloom_doy~threshold,data=dc_g3[-2,])
```



Kyoto, Japan 2023
```{r,echo=T}
kyoto <- read.csv('/Users/nafeezchowdhury/Documents/STAT 490/kyoto.csv',header=T)
kyoto1 <- kyoto %>% filter(1970 <= year & year <= 2000)

kyoto_temp <- ghcnd_search(stationid = "JA000047759",
                     var = c("tmax", "tmin"),
                     date_min = "1970-01-01",
                     date_max = "2000-12-31") %>%
  reduce(left_join) %>%
  transmute(year = parse_number(format(date, "%Y")),
            date, tmax = tmax / 10, tmin = tmin / 10,
            temp = (tmax + tmin) / 2)
kyoto_temp %>% slice(1:4) %>% select(year, date, temp) %>%
  kable(align = "c", col.names =
          c("year", "date", "midrange temperature °C"))

##########

kyoto_year_temps <-
  kyoto_temp %>% 
  mutate(doy_start = parse_number(format(date, "%j"))) %>%
  group_by(year) %>% 
  nest() %>%
  mutate(jan_temp = map(data, ~ mean(.$temp[1:30], na.rm = TRUE)),
         feb_temp = map(data, ~ mean(.$temp[31:60], na.rm = TRUE)),
         mar_temp = map(data, ~ mean(.$temp[61:90], na.rm = TRUE))) %>% 
  unnest(jan_temp) %>% 
  unnest(feb_temp) %>%
  unnest(mar_temp) %>%
  ungroup() %>%
  mutate(`very cold` = jan_temp < 4.6 & feb_temp < 4.6,
         `cold` = jan_temp < 4.6, feb_diff = feb_temp - jan_temp, mar_diff = mar_temp - jan_temp,
         `temp level` = factor(c(ifelse(`cold` == T, "cold", "warm"))), 
         grp = factor(c(ifelse(mar_diff < 2.5, "g1",ifelse(mar_diff < 4.0,'g2','g3')))))

##########

kyoto_sum_temps <- kyoto_temp %>% na.omit() %>% select(year,date,temp) %>% group_by(year) %>% 
  filter(1 <= yday(date) & yday(date) <= 100) %>%
  mutate(threshold = cumsum(temp))
kyoto_sum_temps1 <- kyoto_sum_temps %>% group_by(year) %>% nest() %>% left_join(kyoto1) %>%
  ungroup()

vline_data1 <- data.frame(year = kyoto_sum_temps1$year, vline = kyoto_sum_temps1$bloom_doy)
curve_data1 <- kyoto_sum_temps1 %>% pull(data) %>% reduce(bind_rows) %>% mutate(year = format(date, "%Y"))

kyoto_csum_plot <- ggplot() + 
  geom_smooth(aes(x=yday(date),y=threshold,color=), data = curve_data1) +
  geom_vline(aes(xintercept = vline), data = vline_data1, linetype = 2) +
  facet_wrap(~year) + 
  theme_bw() +
  geom_smooth(method = "lm", aes(x=yday(date),y=threshold), 
              data = curve_data1 %>% mutate(doy = parse_number(format(date, "%j"))) %>% filter(doy >= 75), 
              color = "red") + coord_cartesian(ylim=c(0,800))
kyoto_csum_plot

curve_data1$doy <- parse_number(format(curve_data1$date, "%j")) - 75
reg_line <- lm(threshold ~ yday(date):year, curve_data1, subset = doy >= 0)$coefficients
threshold_100_days1 <- reg_line[1] + reg_line[-1]*100

kyoto2 <- kyoto1 %>% mutate(threshold = threshold_100_days1)
kyoto2$`temp level` <- kyoto_year_temps$`temp level`
mod <- lm(kyoto2$bloom_doy~threshold_100_days1)
summary(mod)
kyotoplot <- ggplot(kyoto2,aes(x=threshold,y=bloom_doy,color=`temp level`)) +
  geom_point() + theme_bw() +
  geom_text(aes(label=year),hjust=0, vjust=0)
kyotoplot


ggplot(kyoto_year_temps,aes(x=feb_diff,y=mar_diff,color=grp)) + geom_point() + 
  geom_text(aes(label=year,hjust=-0.5, vjust=0)) +
  geom_text(aes(label=kyoto2$bloom_doy),hjust=1.3, vjust=0) +
  geom_text(aes(label=signif(kyoto2$threshold,3)),hjust=1.3, vjust=1.3)

kyoto_g1 <- kyoto_year_temps %>% mutate(bloom_doy = kyoto2$bloom_doy, threshold = kyoto2$threshold) %>% 
  filter(grp == 'g1')
kyoto_g2 <- kyoto_year_temps %>% mutate(bloom_doy = kyoto2$bloom_doy, threshold = kyoto2$threshold) %>% 
  filter(grp == 'g2')
kyoto_g3 <- kyoto_year_temps %>% mutate(bloom_doy = kyoto2$bloom_doy, threshold = kyoto2$threshold) %>% 
  filter(grp == 'g3')

kyoto2 %>% ggplot(aes(x=threshold,y=bloom_doy,color=kyoto_year_temps$grp)) + 
  geom_smooth(method='lm') + labs(y='bloom day of year',color='Group')

kyoto_temp_g1 <- lm(bloom_doy~threshold,data=kyoto_g1)
summary(kyoto_temp_g1)
plot(bloom_doy~threshold,data=kyoto_g1)
kyoto_temp_g2 <- lm(bloom_doy~threshold,data=kyoto_g2)
summary(kyoto_temp_g2)
plot(bloom_doy~threshold,data=kyoto_g2)
kyoto_temp_g3 <- lm(bloom_doy~threshold,data=kyoto_g3)
summary(kyoto_temp_g3)
plot(bloom_doy~threshold,data=kyoto_g3)
```



Liestal, Switzerland 2023
```{r,echo=T}
liestal <- read.csv('/Users/nafeezchowdhury/Documents/STAT 490/liestal.csv',header=T)
liestal1 <- liestal %>% filter(1980 <= year)

liestal_temp <- ghcnd_search(stationid = "SZ000001940",
                           var = c("tmax", "tmin"),
                           date_min = "1980-01-01",
                           date_max = "2022-12-31") %>%
  reduce(left_join) %>%
  transmute(year = parse_number(format(date, "%Y")),
            date, tmax = tmax / 10, tmin = tmin / 10,
            temp = (tmax + tmin) / 2)
liestal_temp %>% slice(1:4) %>% select(year, date, temp) %>%
  kable(align = "c", col.names =
          c("year", "date", "midrange temperature °C"))

##########

liestal_year_temps <-
  liestal_temp %>% 
  mutate(doy_start = parse_number(format(date, "%j"))) %>%
  group_by(year) %>% 
  nest() %>%
  mutate(jan_temp = map(data, ~ mean(.$temp[1:30], na.rm = TRUE)),
         feb_temp = map(data, ~ mean(.$temp[31:60], na.rm = TRUE)),
         mar_temp = map(data, ~ mean(.$temp[61:90], na.rm = TRUE))) %>% 
  unnest(jan_temp) %>% 
  unnest(feb_temp) %>%
  unnest(mar_temp) %>%
  ungroup() %>%
  mutate(`very cold` = jan_temp < 4.6 & feb_temp < 4.6,
         `cold` = jan_temp < 4.6, feb_diff = feb_temp - jan_temp, mar_diff = mar_temp - jan_temp,
         `temp level` = factor(c(ifelse(`cold` == T, "cold", "warm"))), 
         grp = factor(c(ifelse(mar_diff < 4.0, "g1",ifelse(mar_diff < 5.2,'g2','g3')))))

##########

liestal_sum_temps <- liestal_temp %>% na.omit() %>% select(year,date,temp) %>% group_by(year) %>% 
  filter(1 <= yday(date) & yday(date) <= 115) %>%
  mutate(threshold = cumsum(temp))
liestal_sum_temps1 <- liestal_sum_temps %>% group_by(year) %>% nest() %>% left_join(liestal1) %>%
  ungroup()

vline_data3 <- data.frame(year = liestal_sum_temps1$year, vline = liestal_sum_temps1$bloom_doy)
curve_data3 <- liestal_sum_temps1 %>% pull(data) %>% reduce(bind_rows) %>% mutate(year = format(date, "%Y"))

liestal_csum_plot <- ggplot() +
  geom_smooth(aes(x=yday(date),y=threshold), data = curve_data3) +
  geom_vline(aes(xintercept = vline), data = vline_data3, linetype = 2) +
  facet_wrap(~year) + 
  theme_bw() +
  geom_smooth(method = "lm", aes(x=yday(date),y=threshold), 
              data = curve_data3 %>% mutate(doy = parse_number(format(date, "%j"))) %>% filter(doy >= 75), 
              color = "red") + coord_cartesian(xlim=c(0,100))
liestal_csum_plot

curve_data3$doy <- parse_number(format(curve_data3$date, "%j")) - 75
reg_line3 <- lm(threshold ~ yday(date):year, curve_data3, subset = doy >= 0)$coefficients
threshold_100_days3 <- reg_line3[1] + reg_line3[-1]*100

liestal2 <- liestal1 %>% mutate(threshold = threshold_100_days3)
liestal2$`temp level` <- liestal_year_temps$`temp level`

mod <- lm(liestal2$bloom_doy~threshold_100_days3)
summary(mod)
plot(liestal2$bloom_doy~threshold_100_days3)
identify(threshold_100_days3,liestal2$bloom_doy,label=liestal2$year)

ggplot(liestal_year_temps,aes(x=feb_diff,y=mar_diff,color=grp)) + geom_point() + 
  geom_text(aes(label=year,hjust=-0.5, vjust=0)) +
  geom_text(aes(label=liestal2$bloom_doy),hjust=1.3, vjust=0) +
  geom_text(aes(label=signif(liestal2$threshold,3)),hjust=1.3, vjust=1.3)

liestal_g1 <- liestal_year_temps %>% mutate(bloom_doy = liestal2$bloom_doy, threshold = liestal2$threshold) %>% 
  filter(grp == 'g1')
liestal_g2 <- liestal_year_temps %>% mutate(bloom_doy = liestal2$bloom_doy, threshold = liestal2$threshold) %>% 
  filter(grp == 'g2')
liestal_g3 <- liestal_year_temps %>% mutate(bloom_doy = liestal2$bloom_doy, threshold = liestal2$threshold) %>% 
  filter(grp == 'g3')

liestal2 %>% ggplot(aes(x=threshold,y=bloom_doy,color=liestal_year_temps$grp)) + 
  geom_smooth(method='lm') + labs(y='bloom day of year',color='Group')

liestal_temp_g1 <- lm(bloom_doy~threshold,data=liestal_g1)
summary(liestal_temp_g1)
plot(bloom_doy~threshold,data=liestal_g1)
liestal_temp_g2 <- lm(bloom_doy~threshold,data=liestal_g2)
summary(liestal_temp_g2)
plot(bloom_doy~threshold,data=liestal_g2)
liestal_temp_g3 <- lm(bloom_doy~threshold,data=liestal_g3)
summary(liestal_temp_g3)
plot(bloom_doy~threshold,data=liestal_g3)
```


Prediction

Washington, DC 2023
```{r,echo=T}
jan_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/DC_Jan_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
jan_2023_avg <- ((jan_2023[,1]-32)*5/9 + (jan_2023[,2]-32)*5/9) / 2


feb_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/DC_Feb_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
feb_2023_avg <- ((feb_2023[,1]-32)*5/9 + (feb_2023[,2]-32)*5/9) / 2


mar_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/DC_Mar_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
mar_2023_avg <- ( (mar_2023[,1] - 32) * 5/9 + (mar_2023[,2] - 32) * 5/9) / 2

apr_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/DC_Apr_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:6,37:42),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
apr_2023_avg <- ((apr_2023[,1]-32)*5/9 + (apr_2023[,2]-32)*5/9) / 2

dc_first_4_months <- c(jan_2023_avg,feb_2023_avg,mar_2023_avg,apr_2023_avg) %>% cumsum()
dc_year_2023 <- data.frame(doy = 1:120, threshold = dc_first_4_months)

dc_curve_data_2023 <- dc_year_2023

ggplot() + 
  geom_point(aes(x=doy,y=threshold), data = dc_year_2023 %>% filter(doy < 58), color = "black") + 
  geom_point(aes(x=doy,y=threshold), data = dc_year_2023 %>% filter(doy >= 58), color = "blue") + 
  coord_cartesian(xlim=c(0,100),ylim=c(0,800)) +
  geom_smooth(aes(x=doy,y=threshold), data=dc_curve_data_2023 %>% filter(doy>=75),
              method='lm',color='red', se = FALSE) +
  labs(x = "number of days since January 1st", y = "cumulative temperatures\nobserved (black) and forecasted (blue)")


dc_feb_diff_2023 <- mean(feb_2023_avg) - mean(jan_2023_avg)
dc_feb_diff_2023
dc_mar_diff_2023 <- mean(mar_2023_avg) - mean(jan_2023_avg)
dc_mar_diff_2023

as.Date(75, origin = "2023-01-01")
```

There is a high threshold at 100 days and the curve is similar in appearance to 1990, so the predicted bloom date for Washington, DC this year is March 17, 2023, at 75 days.


Kyoto, Japan 2023
```{r,echo=T}
jan_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Kyoto_Jan_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
jan_2023_avg <- ((jan_2023[,1]-32)*5/9 + (jan_2023[,2]-32)*5/9) / 2


feb_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Kyoto_Feb_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
feb_2023_avg <- ((feb_2023[,1]-32)*5/9 + (feb_2023[,2]-32)*5/9) / 2


mar_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Kyoto_Mar_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
mar_2023_avg <- ((mar_2023[,1]-32)*5/9 + (mar_2023[,2]-32)*5/9) / 2

apr_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Kyoto_Apr_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:6,37:42),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
apr_2023_avg <- ((apr_2023[,1]-32)*5/9 + (apr_2023[,2]-32)*5/9) / 2

kyoto_first_4_months <- c(jan_2023_avg,feb_2023_avg,mar_2023_avg,apr_2023_avg) %>% cumsum()
kyoto_year_2023 <- data.frame(doy = 1:120, threshold = kyoto_first_4_months)

kyoto_curve_data_2023 <- kyoto_year_2023

ggplot() + 
  geom_point(aes(x=doy,y=threshold), data = kyoto_year_2023 %>% filter(doy < 58), color = "black") + 
  geom_point(aes(x=doy,y=threshold), data = kyoto_year_2023 %>% filter(doy >= 58), color = "blue") + 
  coord_cartesian(xlim=c(0,100),ylim=c(0,800)) +
  geom_smooth(aes(x=doy,y=threshold), data=kyoto_curve_data_2023 %>% filter(doy>=75),
              method='lm',color='red', se = FALSE) +
  labs(x = "number of days since January 1st", y = "cumulative temperatures\nobserved (black) and forecasted (blue)")


kyoto_feb_diff_2023 <- mean(feb_2023_avg) - mean(jan_2023_avg)
kyoto_feb_diff_2023
kyoto_mar_diff_2023 <- mean(mar_2023_avg) - mean(jan_2023_avg) 
kyoto_mar_diff_2023

as.Date(98, origin = "2023-01-01")
```

There is a high threshold at 100 days and the curve is similar in appearance to 1972,1973, and 1979, so the predicted bloom date for Kyoto, Japan this year is April 9, 2023, at 98 days.


Liestal, Switzerland 2023
```{r,echo=T}
jan_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Liestal_Jan_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
jan_2023_avg <- ((jan_2023[,1]-32)*5/9 + (jan_2023[,2]-32)*5/9) / 2


feb_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Liestal_Feb_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
feb_2023_avg <- ((feb_2023[,1]-32)*5/9 + (feb_2023[,2]-32)*5/9) / 2


mar_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Liestal_Mar_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
mar_2023_avg <- ((mar_2023[,1]-32)*5/9 + (mar_2023[,2]-32)*5/9) / 2

apr_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Liestal_Apr_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:6,37:42),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
apr_2023_avg <- ((apr_2023[,1]-32)*5/9 + (apr_2023[,2]-32)*5/9) / 2

liestal_first_4_months <- c(jan_2023_avg,feb_2023_avg,mar_2023_avg,apr_2023_avg) %>% cumsum()
liestal_year_2023 <- data.frame(doy = 1:120, threshold = liestal_first_4_months)

liestal_curve_data_2023 <- liestal_year_2023

ggplot() + 
  geom_point(aes(x=doy,y=threshold), data = liestal_year_2023 %>% filter(doy < 58), color = "black") + 
  geom_point(aes(x=doy,y=threshold), data = liestal_year_2023 %>% filter(doy >= 58), color = "blue") + 
  coord_cartesian(xlim=c(0,100),ylim=c(0,800)) +
  geom_smooth(aes(x=doy,y=threshold), data=liestal_curve_data_2023 %>% filter(doy>=75),
              method='lm',color='red', se = FALSE) +
  labs(x = "number of days since January 1st", y = "cumulative temperatures\nobserved (black) and forecasted (blue)")

liestal_feb_diff_2023 <- mean(feb_2023_avg) - mean(jan_2023_avg)
liestal_feb_diff_2023
liestal_mar_diff_2023 <- mean(mar_2023_avg) - mean(jan_2023_avg) 
liestal_mar_diff_2023

as.Date(95, origin = "2023-01-01")
```

There is a moderate threshold at 100 days and the curve is similar in appearance to 1995, 1999, and 2016, so the predicted bloom date for Liestal, Switzerland this year is April 6, 2023, at 95 days.

Vancouver, Canada 2023
```{r,echo=T}
jan_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Vancouver_Jan_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
jan_2023_avg <- ((jan_2023[,1]-32)*5/9 + (jan_2023[,2]-32)*5/9) / 2


feb_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Vancouver_Feb_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,32:35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
feb_2023_avg <- ((feb_2023[,1]-32)*5/9 + (feb_2023[,2]-32)*5/9) / 2


mar_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Vancouver_Mar_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:3,35),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
mar_2023_avg <- ((mar_2023[,1]-32)*5/9 + (mar_2023[,2]-32)*5/9) / 2

apr_2023 <- read_html('/Users/nafeezchowdhury/Documents/STAT 490/Vancouver_Apr_Temps_2023.html') %>%
  html_nodes('div.monthly-calendar') %>%
  html_text2() %>% str_split('° ') %>% unlist() %>% str_split(' ') %>% unlist() %>%
  matrix(ncol=3,byrow=T) %>% .[,-1] %>% .[-c(1:6,37:42),] %>% as.numeric() %>% 
  matrix(ncol=2,byrow=F)
apr_2023_avg <- ((apr_2023[,1]-32)*5/9 + (apr_2023[,2]-32)*5/9) / 2

vancouver_first_4_months <- c(jan_2023_avg,feb_2023_avg,mar_2023_avg,apr_2023_avg) %>% cumsum()
vancouver_year_2023 <- data.frame(doy = 1:120, threshold = vancouver_first_4_months)

vancouver_curve_data_2023 <- vancouver_year_2023

ggplot() + 
  geom_point(aes(x=doy,y=threshold), data = vancouver_year_2023 %>% filter(doy < 58), color = "black") + 
  geom_point(aes(x=doy,y=threshold), data = vancouver_year_2023 %>% filter(doy >= 58), color = "blue") + 
  coord_cartesian(xlim=c(0,100),ylim=c(0,800)) +
  geom_smooth(aes(x=doy,y=threshold), data=vancouver_curve_data_2023 %>% filter(doy>=75),
              method='lm',color='red', se = FALSE) +
  labs(x = "number of days since January 1st", y = "cumulative temperatures\nobserved (black) and forecasted (blue)")

vancouver_feb_diff_2023 <- mean(feb_2023_avg) - mean(jan_2023_avg)
vancouver_feb_diff_2023
vancouver_mar_diff_2023 <- mean(mar_2023_avg) - mean(jan_2023_avg) 
vancouver_mar_diff_2023

as.Date(98, origin = "2023-01-01")

```

The predicted bloom date for Vancouver is April 9, 2023 at 98 days.

